### 2018-02-23 paper/sections/1_introduction.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

ParSplice molecular dynamics simulations are representative of an important
class of HPC applications with similar working set behaviors that extensively
use software-based caches. ParSplice uses a hierarchy of caches and a single
persistent key-value store to store the atomic coordinates corresponding
to local energy minima (referred to simply as minima) encountered
by a large number of independent dynamical trajectories.
This workload is also pervasive
across simulations that (1) rely on a mesh-based decomposition of a physical
region and (2) result in millions or billions of mesh cells, where each cell
contains materials, pressures, temperatures and other characteristics that are
required to accurately simulate phenomena of interest.  The fine-grained data
annotation capabilities provided by key-value storage is a natural match for
these types of scientific simulations.  Unfortunately, simulations of this size
saturate the capacity and bandwidth capabilities of a single node so we need
more effective data management techniques.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
Solutions labeled ``no performance degradation" are in comparison to the
baseline of using a cache of unlimited size (``No Cache Management").
\label{fig:cache-management}}
\end{figure}

A challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores atomic
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, this threshold works well
for HPC applications.  For reference, a 1GB cache for a distributed file system
is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
300 worker processes, so large simulations need more caches and will use even
more memory.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the cost of performance, which
is shown by the text annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without any performance degradation, works for a variety of initial conditions,
and could generalize to similar applications. Triggering key eviction at a
certain memory pressure ({\it e.g.} 3\%) requires {\it a priori} system
knowledge while our approach saves the most memory because we model the
behavior of keyspace accesses.

>>>>

ParSplice molecular dynamics simulations are representative of an important
class of HPC applications with similar working set behaviors that extensively
use software-based caches. It uses a hierarchy of caches and a single
persistent key-value store to store the atomic coordinates corresponding
to local energy minima (referred to simply as minima) encountered
by a large number of independent dynamical trajectories.
This workload is also pervasive
across simulations that (1) rely on a mesh-based decomposition of a physical
region and (2) result in millions or billions of mesh cells, where each cell
contains materials, pressures, temperatures and other characteristics that are
required to accurately simulate phenomena of interest.  The fine-grained data
annotation capabilities provided by key-value storage is a natural match for
these types of scientific simulations.  Unfortunately, simulations of this size
saturate the capacity and bandwidth capabilities of a single node so we need
more effective data management techniques.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
Solutions labeled ``no performance degradation" are in comparison to the
baseline of a cache of unlimited size.
\label{fig:cache-management}}
\end{figure}

A challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores atomic
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, this threshold works well
for HPC applications.  For reference, a 1GB cache for a distributed file system
is too large in LANL deployments.}. Furthermore, ParSplice deploys a cache per
300 worker processes, so large simulations need more caches and will use even
more memory.  Users can configure ParSplice to evict data when the cache
reaches a threshold but this solution requires tuning and parameter sweeps; the
``Cache (too small)" curve in Figure~\ref{fig:cache-management} shows how a
poorly configured cache can save memory but at the cost of performance, which
is shown by the text annotation to the right.  Even worse, this threshold
changes with different initial configurations and cluster setups so tuning
needs to be done for all system permutations.  Our dynamically sized cache,
shown by the thick line in Figure~\ref{fig:cache-management}, detects key
access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache
without any performance degradation, works for a variety of initial conditions,
and could generalize to similar applications. Triggering key eviction at a
certain memory pressure ({\it e.g.}, 3\%) requires {\it a priori} system
knowledge while our approach saves the most memory because we model the
behavior of keyspace accesses.

### 2018-02-23 paper/sections/2_parsplice-keyspace-analysis.tex
paper: parsplice corrections from danny
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsubsection{\underline{Initial conditions influence key activity}}
\label{sec:delta}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The access patterns for
different growth rates have temporal locality, as the reads per second for
\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
microseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
microseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
a temperature of 400 degrees result in very little database activity because
state transitions take too long. Similarly, values larger than \(\Delta_1\)'s
growth rate or a temperature of 4000 degrees result in an equally meaningless
simulation as transitions are unrealistic.

This figure demonstrates that small changes to \(\Delta\) can have a strong
effect on the timing and frequency with which new EOM minima are discovered and
referenced.  Trends also exist for temperature and number of workers but are
omitted here for space.  This finding suggests that we need a flexible policy
language and engine to explore these trade-offs.

>>>>

\subsubsection{\underline{Initial conditions influence key activity}}
\label{sec:delta}
Figure~\ref{fig:motivation-regimes} shows how ParSplice tasks read key-value
pairs from the worker's cache for two different initial conditions of
\(\Delta\), which is the rate that new atoms enter the simulation.  The line is
the read request rate (\(y1\) axis) and the dots along the bottom are the
number of unique keys accessed (\(y2\) axis).  The access patterns for
different growth rates have temporal locality, as the reads per second for
\(\Delta_2\) look like the reads per second for \(\Delta_1\) stretched out
along the time axis.  The \(\Delta_1\) growth rate adds atoms every 100K
picoseconds while the \(\Delta_2\) growth rate adds atoms every 1 million
picoseconds. So \(\Delta_2\) has a smaller growth rate resulting in hotter
keys and a smaller keyspace.  Values smaller than \(\Delta_2\)'s growth rate or
a temperature of 400 degrees result in very little database activity because
state transitions take very long. Similarly, growth rate values larger than \(\Delta_1\)'s
lead to far out-of-equilibrium growth, while temperatures in excess of 800 degrees
result in the melting of the nanoparticle, which is not relevant in this application.

This figure demonstrates that small changes to \(\Delta\) can have a strong
effect on the timing and frequency with which new minima are discovered and
referenced.  Trends also exist for temperature and number of workers but are
omitted here for space.  This finding suggests that we need a flexible policy
language and engine to explore these trade-offs.
